### 2016-08-08 Projects/DelCo_2014_2015/Reports/Hose_Stream_Tests/TeX_Files/DelCo_Hose_Stream_Tests.tex
update Hosestream report from JH comments

\begin{tabbing}
\hspace{1.5in} \= \\
ESTC \> Delaware County Emergency Services Training Center \\
NIST \> National Institute of Standards and Technology \\
UL FSRI \> Underwriters Laboratories Firefighter Safety Research Institute
\end{tabbing}

@@

Fire suppression tactics using hose streams can affect ventilation in a structure and may impact the movement of smoke and heat through a structure. Seven series of experiments were conducted to study the impact of different hose stream and nozzle movement pattern combinations on air movement within residential scale structures. The experiments studied four different hose streams: a straight stream, narrow fog stream, and a wide fog stream from a combination nozzle and a solid stream from a smooth bore nozzle. The streams were applied from a static, or fixed, position; by moving the hoseline left-to-right across the room in a sweeping motion; and by rotating the hoseline in both the clockwise and counterclockwise directions. Gas velocity was measured at different locations in the structure during the experiments. The wide fog stream caused the most air movement out of any of the tested streams, reaching a maximum velocity of 2.6~m/s (5.8~mph) and maximum air flow rate of approximately 4.2~m$^3$/s (9000~cfm), followed by the narrow fog stream, and then the straight stream and solid stream from the smooth bore nozzle, which both caused approximately the same amount of air movement through the structure. The data were consistent with a straight stream only causing air movement through the structure when it was applied in a moving pattern. Furthermore, there were no statistically significant differences in the average measured air velocity between the clockwise and counterclockwise nozzle movement patterns. It was determined that the type of hose stream and manner in which it is applied dictates the extent to which a stream impacts the ventilation of a structure. To better quantify the impact of such air flows on the fire environment, additional experiments need to be conducted with structure fires in a controlled environment.

@@

The first floor ceiling support of each structure was composed of wood truss joist I-beams (TJIs) with a 298~mm (11.75~in) depth. Each TJI was composed of laminated veneer lumber flanges with a cross section of 29~mm (1.13~in) x 44~mm (1.75~in) and an 11~mm (0.43~in) thick oriented strand board web as shown in Fig.~\ref{fig:TJI}. Tongue and groove oriented strand board of 18.3~mm (0.72~in) thickness was screwed to the top of the TJIs.

@@

\subsection{Uncertainty}
\label{sec:uncertainty}
There are different components of uncertainty in the length, differential pressure, flow rate, and gas velocity reported here. Uncertainties are grouped into two categories according to the method used to estimate them. Type A uncertainties are those which are evaluated by statistical methods, and Type B are those which are evaluated by other means~\cite{Taylor&Kuyatt:1994}. Type B analysis of systematic uncertainties involves estimating the upper (+a) and lower (-a) limits for the quantity in question such that the probability that the value would be in the interval ($\pm$a) is essentially 100~\%. After estimating uncertainties by either Type A or B analysis, the uncertainties are combined in quadrature to yield the combined standard uncertainty. Then, the combined standard uncertainty is multiplied by a coverage factor of two, which results in the expanded uncertainty with a 95~\% confidence interval (2$\sigma$). For some of these components, such as the zero and calibration elements, uncertainties are derived from referenced instrument specifications. For other components, referenced research results and past experience with the instruments provided input in the uncertainty determination.

@@

\section{Experimental Procedure}
\label{sec:exp_procedure}
Every experiment discussed in this report involved discharging water at a set flow rate from a combination nozzle in a straight, narrow fog, or wide fog stream pattern or from a smooth bore nozzle with a 1~in tip in a solid stream pattern. Specific doors were opened or closed before and/or during the tests to change the ventilation within the structure.

@@

% ==========================
% = RESULTS AND DISCUSSION =
% ==========================
\chapter{Results and Discussion}
\label{chap:results}
In the following sections, the measurements are presented in graphic and tabular form. In the graphs, an error bar will represent the estimated uncertainty of the measurement. In the tables, the uncertainty will be included in the caption of the table as a percentage enclosed in brackets.

@@

\multirow{5}{*}{\textit{\small Straight}} & \multirow{2}{*}{Velocity}

@@

\multirow{5}{*}{\begin{tabular}{@{}c@{}} \textit{\small Narrow} \\ \textit{\small Fog} \end{tabular}} & \multirow{2}{*}{Velocity}

@@

\multirow{5}{*}{\begin{tabular}{@{}c@{}} \textit{\small Wide} \\ \textit{\small Fog} \end{tabular}} & 

@@

\multirow{5}{*}{\begin{tabular}{@{}c@{}} \textit{\small Smooth} \\ \textit{\small Bore} \end{tabular}} &

@@

Additionally, a fourth experimental series using a monitor was conducted in which a straight stream from a combination nozzle was compared to a solid stream from a smooth bore nozzle with a 1~in tip aimed at a specified location for 60 seconds with a fully established flow path. The streams were compared for six different 60 seconds periods during which the flow path was fully established: three times each for the monitor aimed at two locations. The results from the four monitor test series suggest a wide fog stream impacts air movement within a structure the most out of any of the tested streams, followed by the narrow fog stream, and then the straight stream and solid stream from a smooth bore nozzle, which both caused approximately the same amount of air movement through the structure.

>>>>

\begin{tabbing}
\hspace{1.5in} \= \\
ESTC \> Delaware County Emergency Services Training Center \\
NF \> Narrow Fog Stream\\
NIST \> National Institute of Standards and Technology \\
SB \> Smooth Bore Stream\\
SS \> Straight Stream \\
UL FSRI \> Underwriters Laboratories Firefighter Safety Research Institute
WF \> Wide Fog Stream \\
\end{tabbing}

@@

Fire suppression tactics using hose streams can affect ventilation in a structure and may impact the movement of smoke and heat through a structure. Seven series of experiments were conducted to study the impact of different hose stream and nozzle movement pattern combinations on air movement within residential scale structures. The experiments studied four different hose streams: a straight stream, narrow fog stream, a wide fog stream from a combination nozzle, and a solid stream from a smooth bore nozzle. The streams were applied from a static, or fixed, position; by moving the hoseline left-to-right across the room in a sweeping motion; and by rotating the hoseline in both the clockwise and counterclockwise directions. Gas velocity was measured at different locations in the structure during the experiments. The wide fog stream caused the most air movement out of any of the tested streams, reaching a maximum velocity of 2.6~m/s (5.8~mph) and maximum air flow rate of approximately 4.2~m$^3$/s (9000~cfm), followed by the narrow fog stream, and then the straight stream and solid stream from the smooth bore nozzle, which both caused approximately the same amount of air movement through the structure. The data were consistent with a straight stream only causing air movement through the structure when it was applied in a moving pattern. Furthermore, there were no statistically significant differences in the average measured air velocity between the clockwise and counterclockwise nozzle movement patterns. It was determined that the type of hose stream and manner in which it is applied dictates the extent to which a stream impacts the ventilation of a structure. To better quantify the impact of such air flows on the fire environment, additional experiments need to be conducted with structure fires in a controlled environment.

@@

The first floor ceiling support of each structure was composed of wood truss joist I-beams (TJIs) with a 298~mm (11.75~in) depth. Each TJI was composed of laminated veneer lumber flanges with a cross section of 29~mm (1.13~in) x 44~mm (1.75~in) and an 11~mm (0.43~in) thick oriented strand board (OSB) web as shown in Fig.~\ref{fig:TJI}. Tongue and groove oriented strand board of 18.3~mm (0.72~in) thickness was screwed to the top of the TJIs.

@@

\subsection{Uncertainty}
\label{sec:uncertainty}
There are different components of uncertainty in the length, differential pressure, flow rate, and gas velocity presented in this report. Uncertainties are grouped into two categories according to the method used to estimate them. Type A uncertainties are those which are evaluated by statistical methods, and Type B are those which are evaluated by other means~\cite{Taylor&Kuyatt:1994}. Type B analysis of systematic uncertainties involves estimating the upper (+a) and lower (-a) limits for the quantity in question such that the probability that the value would be in the interval ($\pm$a) is essentially 100~\%. After estimating uncertainties by either Type A or B analysis, the uncertainties are combined in quadrature to yield the combined standard uncertainty. Then, the combined standard uncertainty is multiplied by a coverage factor of two, which results in the expanded uncertainty with a 95~\% confidence interval (2$\sigma$). For some of these components, such as the zero and calibration elements, uncertainties are derived from referenced instrument specifications. For other components, referenced research results and past experience with the instruments provided input in the uncertainty determination.

@@

\section{Experimental Procedure}
\label{sec:exp_procedure}
Every experiment discussed in this report involved discharging water at a set flow rate from a combination nozzle in a straight, narrow fog, or wide fog stream pattern or from a smooth bore nozzle with a 1~in tip with a solid stream pattern. Specific doors were opened or closed before and/or during the tests to change the ventilation within the structure.

@@

% ==========================
% = RESULTS AND DISCUSSION =
% ==========================
\chapter{Results and Discussion}
\label{chap:results}
In the following sections, the measurements are presented in graphic and tabular form. In the graphs, an error bar will represent the estimated uncertainty of the measurement. In the tables, the uncertainty will be included in the caption of each table as a percentage enclosed in brackets.

@@

\multirow{5}{*}{\textit{\small Straight (SS)}} & \multirow{2}{*}{Velocity}

@@

\multirow{5}{*}{\begin{tabular}{@{}c@{}} \textit{\small Narrow} \\ \textit{\small Fog (NF)} \end{tabular}} & \multirow{2}{*}{Velocity}

@@

\multirow{5}{*}{\begin{tabular}{@{}c@{}} \textit{\small Wide} \\ \textit{\small Fog (WF)} \end{tabular}} & 

@@

\multirow{5}{*}{\begin{tabular}{@{}c@{}} \textit{\small Smooth} \\ \textit{\small Bore (SB)} \end{tabular}} &

@@

Additionally, a fourth experimental series using a monitor was conducted in which a straight stream from a combination nozzle was compared to a solid stream from a smooth bore nozzle with a 1~in tip aimed at a specified location for 60 seconds with a fully established flow path. The streams were compared for six different 60 second periods during which the flow path was fully established: three times each for the monitor aimed at two locations. The results from the four monitor test series suggest a wide fog stream impacts air movement within a structure the most out of any of the tested streams, followed by the narrow fog stream, and then the straight stream and solid stream from a smooth bore nozzle, which both caused approximately the same amount of air movement through the structure.
